Assessing Tor's Usability as a Censorship Circumvention Tool

### 2015-05-12 cs261n-paper/paper.tex
reword related work
Edits on pilot study.

Tor Browser~\cite{tor-browser},
a specially modified version of Firefox that comes
with a built-in Tor proxy and pluggable transports,
is the primary means through which users access the Tor network.
It aims to make powerful anonymity and circumvention technology
accessible even to nontechnical users.
It has not undergone a formal usability assessment since
its user interface was overhauled in its version~3.5 release in December 2013.
A central component of its user configuration interface
is a panel to select bridges and pluggable transports
for censorship circumvention.
This selection interface is the focus of our study.

@@

\section{A Pilot Tor Usability Study} 
\indent \indent We ran a pilot Tor usability study to test the usability of Tor as an anonymity tool,
targeting a specific demographic of users known to use Tor (journalists who have reported in other
countries) and aiming to identify how well they could understand Tor's non-typical behaviors and 
navigate advanced settings. For the full details of this pilot experiment, we refer you to Appendix~\ref{sec:pilot}.

During this pilot, the users were struggling with confusing bugs, such as flash errors which
say ``Firefox'' rather than ``Tor,'' disappearing buttons on the configuration window, and errors when
launching Tor from the download dialog rather than from their applications folder. This resulted in our 
participants spending most of their time on the download and install tasks, rather than 
our original task for navigating a robust menu or questioning the clarity of provided security 
guarantees. We took the opportunity to collaborate with Tor developers to make changes for the Tor 
Browser and use the lessons learned to design our large-scale user study on testing Tor as a 
censorship circumvention tool.

\subsection{Impact}
\indent \indent Along with the Tor developers who came to observe the UX pilot, we filed 
a total of sixteen bug tickets to Tor. At the time of  this writing, eight of the sixteen tickets 
have already been resolved, and more will be resolved in the future.  The list of bug
tickets filed can be found at \url{https://trac.torproject.org/projects/tor/query?keywords=~uxsprint2015}.

As of April, the changes resulting from our pilot study has been incorporated into the latest 
version of Tor Browser, making Tor more usable for people to use. 
The summary of the usability changes can be found at \url{https://blog.torproject.org/blog/tor-browser-45a4-released}.

Through the publicity of this usability pilot study on the Tor weekly blogs, we were connected
with a source from the Library Freedom Project, which provides training on various technologies 
to libraries across the country. We now collaborate with her work on creating a user manual
for the Tor Browser.

@@

Although the data from the pilot study provided a lot of insight into what users thought processes
are when completing various asks, transcribing audio recordings and syncing them to videos was
required a lot of manual labor. Not only did this form of data collection and processing not scale 
to larger numbers, we found that most users encountered similar thought processes and struggled
with the same issues. Because of this, we aim to collect our future qualitative data in the form of 
an open-ended survey question rather than using audio recordings.

@@

\section{UX sprint}
\label{sec:pilot}
\indent \indent This pilot was to test-drive an experiment which examines the usability of Tor as an anonymity tool
with respect to a target demographic of users known to use Tor (journalists who have reported in other
countries). We emphasized the targeting of a high-risk user group of Tor and to 
target usability changes aimed at helping those who could not afford to make mistakes. To see 
if those users had any potentially damaging misunderstandings or incomprehensions of Tor, 
we created tasks to elicit Tor's non-typical behaviors and ask our participants to
navigate through more advanced settings.

\subsection{Experiment Overview}
\indent \indent Our experiment consisted of three parts. We first prepare our participants to perform a 
cognitive walkthrough, ask our participants to perform a list of Internet browsing 
tasks while performing a cognitive walkthrough, and finish with an exit survey to collect 
additional information.

To first ensure that our users knew what a cognitive walkthrough is, we explain what a cognitive
walkthrough and give a demonstration of a cognitive walkthrough on a non-related dummy task 
(drawing a red hexagon in the computer graphics program Paint). We then ask our users to perform
an unrelated Internet-based task (finding the population of Zimbabwe) using any browser of their choice
while performing a cognitive walkthrough along the process of completing the task. We give our 
participants feedback on how to properly conduct a cognitive walkthrough~\cite{wharton1994cognitive}.
This process was also used to get users comfortable with talking our loud with the researcher and gauge 
how comfortable users were with a computer.

@@

We originally had another step after Step~1, ``Configure Tor.''
We removed it after the first two participants
when it became clear that the wording was confusing.
The two main buttons in the configuration wizard are labeled
``Connect'' and ``Configure''; we did not intend for users
to have to go through the longer ``Configure'' path,
but they assumed they had to because of the presence
of the word ``Configure'' in the list of tasks.

Tasks 1--3 test how easy it was for a user to search for, download, and install Tor Browser 
correctly for use and if they understood that they had done the tasks correctly. 
Since no software is useful if the user is not able to access it, we found it valuable
to test the usability of this process.  Additionally, tasks 4--6 highlight Tor's ``irregular'' behavior 
comparatively to other web browsers. Specifically, gauge user reactions to Tor's default
privacy-preserving search engine Startpage,  longer loading times and performance for videos, 
and disappearing tabs upon starting a new identity. Lastly, we asked our users to 
interact with and describe buttons on the toolbar to gauge user comprehension of how Tor functioned.

@@

\subsection{Methodology} 
\indent \indent We recruited five journalists from Berkeley by reaching out to journalist contacts.
 We recruited five users with an average age of 28 ($\sigma = 13 $~years). Out of 
our five participants, 3 were female.

@@

This pilot was held in conjunction with the first Tor UX sprint. With IRB approval, Tor developers
observed the screens of our participants, which was set to broadcast to a separate  
room during the experiment. Audio of the cognitive walkthrough was not broadcast for privacy
considerations as voice is considered an identifying characteristic.

\subsection{Results}
\indent \indent This experiment uncovered a powerful vector for attack, an unlikely source
of usability issues, and insight into how to better communicate to users so that they can be 
empowered to use Tor Browser and its advanced settings.

\subsubsection{How to Attack Users}
During the task completion process, we observed that all our participants are likely 
susceptible to a search engine optimization (SEO) poisoning attack. Most of our 
participants  were on autopilot to complete this standard download task and 
were also not aware of the possibility that their download may come from an 
illegitimate source. Usually, our participants picked whichever was the first search
result and believed it to be a good enough source for the download:

@@

Luckily, all of our users managed to download Tor browser from the legitimate source. However, 
upon searching through various search engines with a clear history, we found that cnet.com or techsource.com 
 appears as the first search result. Defenses against SEO optimization would be prudent to employ.

\subsubsection{Fighting Mac's Security Features}
The biggest usability obstacle was the result of tensions between Mac's Gatekeeper 
code-signing feature on OS X, which prevents users from opening unsigned code 
downloaded from the Internet to execute on their machine unless overridden through
managing system preferences. Surprisingly, the feature which impacted Tor's usability
the most was the tension between a security precaution, and Tor's decision to self-sign
their downloads, rather than a specific feature of the Tor Browser itself. We found that 
encountering the Gatekeeper will cause frustration to most users, and may be enough 
to stop our participants in the download process:

@@

Notice that this user explained what the feature can do for him, rather than how the 
feature works. For the end user, the reason why a security feature is employed is what 
makes a feature valuable and thus motivates them to use the feature. Communicating why a 
feature is present may be more effective for increasing use of advanced features 
than communicating the technical details of how a feature works.

@@

\indent \indent Because of the limited size of the study, it's not possible to state with confidence what 
fraction of users will encounter major usability obstacles. However, it was effective at discovering 
and demonstrating issues that are likely to cause problems for many users. 
People did have difficulty with installing Tor Browser (principally because of the Gatekeeper 
code-signing feature on OS X), did not understand what many of the toolbar options meant, 
and were confused by most Tor-specific behaviors. For specific problems 
encountered by an individual user, we refer you to our resources online.

>>>>

Tor Browser~\cite{tor-browser},
a specially modified version of Firefox that comes
with a built-in Tor proxy and pluggable transports,
is the primary means through which users access the Tor network.
It aims to make powerful anonymity and circumvention technology
accessible even to nontechnical users.
It has not undergone a formal usability assessment since
its user interface was overhauled in December 2013.
A central component of its user configuration interface
is a panel to select bridges and pluggable transports
for censorship circumvention.
This selection interface is the focus of our study.

@@

\section{A Pilot Tor Usability Study} 
\indent \indent We ran a pilot study to test the usability of Tor as an anonymity tool,
targeting a demographic of users who may have reason to use Tor, namely, journalists who have reported in other
countries. We sought to identify how well they could understand Tor's non-typical behaviors and 
navigate advanced settings. For the full details of this pilot experiment, see Appendix~\ref{sec:pilot}.

During the pilot, users struggled with confusing bugs, such as a few error messages that
said ``Firefox'' rather than ``Tor Browser,'' disappearing buttons on the configuration window, and errors when
running the program from the download archive without installing it.
This resulted in our
participants spending most of their time on the download and install tasks, and not
what we considered the main tasks of browsing and
questioning the clarity of provided security 
guarantees. We took the opportunity to collaborate with Tor developers to make changes in the
browser and use the lessons learned to design our large-scale user study on testing Tor as a 
censorship circumvention tool.

\subsection{Impact}
\indent \indent Along with the Tor developers who came to observe the usability pilot, we filed 
a total of sixteen bug tickets to Tor. At the time of  this writing, eight of the sixteen tickets 
have already been resolved, and more will be resolved in the future.  The list of bug
tickets filed can be found at \url{https://trac.torproject.org/projects/tor/query?keywords=~uxsprint2015}.

As of April 2015, the usability improvements resulting from our pilot study has been incorporated into the shipping
release of Tor Browser.
A summary of changes can be found at \url{https://blog.torproject.org/blog/tor-browser-45a4-released}.

Through the publicity of the usability study on the Tor blog, we were connected
with a source from the Library Freedom Project~\cite{libraryfreedomproject}, which provides training on various technologies 
to libraries across the country. We now collaborate with her in working on creating a user manual
for Tor Browser.

@@

Although the data from the pilot study provided a lot of insight into what users thought processes
are when completing various asks, transcribing audio recordings and syncing them to videos
required much manual work. Not only did this form of data collection and processing not scale 
to larger numbers, we found that most users encountered similar thought processes and struggled
with the same issues. Because of this, we aim to collect our future qualitative data in the form of 
an open-ended survey rather than through audio recordings.

@@

\section{Pilot study}
\label{sec:pilot}
\indent \indent This pilot examined the usability of Tor as an anonymity tool
with respect to journalists, a target demographic of users known to use Tor. 
We emphasized the targeting of a high-risk user group of Tor and to 
target usability changes aimed at helping those who could not afford to make mistakes. To see 
if those users had any potentially damaging misunderstandings or incomprehensions of Tor, 
we created tasks to elicit Tor's non-typical behaviors and ask our participants to
navigate through more advanced settings.

\subsection{Experiment Overview}
\indent \indent Our experiment consisted of three parts. We first prepared our participants to perform a 
cognitive walkthrough, asked our participants to perform a list of Internet browsing 
tasks while performing a cognitive walkthrough, and finished with an exit survey to collect 
demographic information.

To first ensure that our participants are able to perform a cognitive walkthrough is, we explain what a cognitive
walkthrough is and gave a demonstration of a cognitive walkthrough on a non-related dummy task 
(drawing a red hexagon in the computer graphics program Paint). We then asked our users to perform
an unrelated Internet-based task (finding the population of Zimbabwe) using any browser of their choice
while performing a cognitive walkthrough. We took this opportunity to give our 
participants feedback on how to properly conduct a cognitive walkthrough~\cite{wharton1994cognitive}.
This process was also used to get users comfortable with talking our loud with the researcher and gauge 
how comfortable users were with a computer.

@@

We originally had an extra step after Step~1: ``Configure Tor.''
We removed it after the first two participants
when it became clear that the wording was confusing.
The two main buttons in the configuration wizard are labeled
``Connect'' and ``Configure''; we did not intend for users
to have to go through the longer ``Configure'' path,
but they assumed they had to because of the presence
of the word ``Configure'' in the list of tasks.

Tasks 1--3 tested how easy it was for a user to search for, download, and install Tor Browser 
correctly for use and if they understood that they had done the tasks correctly. 
Since no software is useful if the user is not able to access it, we found it valuable
to test the usability of this process.  Additionally, tasks 4--6 highlighted Tor's ``irregular'' behavior 
comparatively to other web browsers. Specifically, we wanted to gauge user reactions to Tor's
privacy-preserving search engine Startpage, decreased performance, 
and side effects of advanced settings (such as disappearing tabs upon starting a new identity).
 Lastly, we asked our users to 
interact with and describe buttons on the toolbar to gauge user comprehension of how Tor functioned.

@@

\subsection{Methodology} 
\indent \indent We recruited five journalists from the Berkeley area by reaching out to journalist contacts.
 We recruited five users with an average age of 28 ($\sigma = 13 $~years).  Out of  %?
five participants, three were female.

@@

This pilot was held in conjunction with the first Tor UX Sprint. With IRB approval, Tor developers
observed the screens of our participants, which was set to broadcast to a separate  
room during the experiment. Audio of the cognitive walkthrough was not broadcast for privacy
considerations as voice is considered an identifying characteristic.

\subsection{Results}
\indent \indent This experiment uncovered a powerful vector for attack, an unlikely source
of usability issues, and how users communicated with the current interface design.

\subsubsection{How to Attack Users}
During the task completion process, we observed that all our participants are likely 
susceptible to a search engine optimization (SEO) poisoning attack during the 
Tor download process. Most of our  participants were on autopilot and 
were also not aware of the possibility that their download may come from an 
illegitimate source. Usually, our participants picked whichever was the first search
result and believed it to be a good enough source for the download:

@@

Luckily, all of our users managed to download Tor browser from the legitimate source. However, 
defenses against SEO optimization would be prudent to employ.

\subsubsection{Fighting Mac's Security Features}
The biggest usability obstacle was the result of tensions between Mac's Gatekeeper 
code-signing feature on OS X, which prevents users from opening unsigned code 
downloaded from the Internet to execute on their machine unless overridden through
managing system preferences. Participants could not download Tor without correctly
configuring these system preferences. We found that 
encountering the Gatekeeper will cause frustration to most users, and may be enough 
to stop our participants in the download process:

@@

This user explained what the feature can do for him, rather than how the 
feature works. For the end user, the reason why a security feature is employed is what 
makes a feature valuable and thus motivates them to use the feature. We believe that communicating why a 
feature is present may be more effective for increasing use of advanced features 
than communicating the technical details of how a feature works.

@@

\indent \indent Because of the limited size of the study, it's not possible to state with confidence what 
fraction of users will encounter major usability obstacles. However, it was effective at discovering 
and demonstrating issues that are likely to cause problems for many users. For specific problems 
encountered by an individual user, we refer you to our resources online.
